---
title: "Q2 HW3"
author: "Rstudio"
date: "2022-12-04"
output: html_document
---


## Part 2: MovieLens Data

1.\ Load in the training dataset from [here](https://github.com/datasciencelabs/2022/blob/master/data/mv_st.RDS), which can be accessed as `mv_st$train`.


```{r, echo=FALSE}
setwd("C:\\Users\\Youval Fouks\\Desktop\\BST219_2022\\fouksiR- 2022HW3")
list.files()
mv1 = readRDS("mv_st.RDS") 

head(mv1)
```

```{r, echo=FALSE}}
if (!require(MASS)) {install.packages("MASS")}
if (!require(caret)) {install.packages("caret")}
if (!require(leaps)) {install.packages("leaps")}
if (!require(gamlr)) {install.packages("gamlr")}
if (!require(glmnet)) {install.packages("glmnet")}
if (!require(purrr)) {install.packages("purrr")}
if (!require(pROC)) {install.packages("pROC")}
if (!require(survivalROC)) {install.packages("survivalROC")}
if (!require(tidyr)) {install.packages("tidyr")}
if (!require(ggplot2)) {install.packages("ggplot2")}
if (!require(arsenal)) {install.packages("arsenal")}
```

2.\ Preprocess the data in any way you find appropriate. This could include removing noninformative features.

```{r, echo=FALSE}}
library(tidyr)
library(dplyr)

names(mv1)

test <- mv1$test
train <- mv1$train

class(mv1$train)
class(mv1$test)

df.train <- as.data.frame(train)
df.test <- as.data.frame(test)
```

```{r, echo=FALSE}}
unique(train$genres)
```
```{r, echo=FALSE}}
library(stringr)
train1 <- separate_rows(df.train, genres, sep='\\|')
test1 <- separate_rows(df.test, genres, sep='\\|')
train1
```

```{r, echo=FALSE}}
naniar::miss_var_summary(train1)
naniar::pct_miss(train1)
```
```{r, echo=FALSE}}
train1 <- train1 |> 
  pivot_wider(names_from = genres, values_from = genres)
select(train1, title, year, userId, rating, timestamp)

test1 <- test1 |> 
  pivot_wider(names_from = genres, values_from = genres)
select(test1, title, year, userId, rating, timestamp)
```
recode var
```{r, echo=FALSE}}

cols.factor <- names(train1)[7:25] # or column index (change the index if needed)
train1[cols.factor] <- lapply(train1[cols.factor], as.factor)

library(dplyr)
train1 <- train1 %>%
  mutate_if(is.factor, funs(ifelse(is.na(.), 0, .)))

cols.factor <- names(test1)[7:25] # or column index (change the index if needed)
test1[cols.factor] <- lapply(test1[cols.factor], as.factor)

library(dplyr)
test1 <- test1 %>%
  mutate_if(is.factor, funs(ifelse(is.na(.), 0, .)))

#replace(is.na(train1), 0)
#train1[train1 == "NA"] <- "0"
#train1$Animation <-ifelse(train1$Animation == "NA" , 0, 1)
```

```{r, echo=FALSE}}
naniar::miss_var_summary(train1)
naniar::pct_miss(train1)
```

Table descriptive
```{r, echo=FALSE}}
cols.factor <- names(train1)[7:25] # or column index (change the index if needed)
train1[cols.factor] <- lapply(train1[cols.factor], as.factor)

cols.factor <- names(test1)[7:25] # or column index (change the index if needed)
test1[cols.factor] <- lapply(test1[cols.factor], as.factor)
```
```{r, echo=FALSE}}
library(arsenal)
Table1 <- tableby(rating ~ title + year + userId + timestamp ,data= train1)
summary(Table1)
```

3.\ Train a machine learning model. Choose at least one model validation method (data splitting, n-fold cross validation, and bootstrapping) that you learned in class to determine how well your model is doing in each case.

Data Partition
```{r, echo=FALSE}}
df <-train1
names(df)[14]=paste("Scif")
names(df)[25]=paste("Noir")

df.t <-test1
names(df.t)[10]=paste("Scif")
names(df.t)[23]=paste("Noir")

df = subset(df, select = -c(2) )
df.t = subset(df.t, select = -c(2) )
```

more manageable subset. We will take the first 1,000 predictors x and labels y:
```{r, echo=FALSE}}
train_subset <- df[1:10000,] 
df <- train_subset

test_subset <- df.t[1:10000,] 
###################
#####################
df.t <- test_subset
```

```{r, echo=FALSE}}
## 75% of the sample size
smp_size <- floor(0.75 * nrow(df))

## set the seed 
set.seed(2050)
train_ind <- sample(seq_len(nrow(df)), size = smp_size)

train <- df[train_ind, ]
test <- df[-train_ind, ]

## 75% of the sample size
smp_size.t <- floor(0.75 * nrow(df.t))

## set the seed 
set.seed(2051)
test_ind <- sample(seq_len(nrow(df.t)), size = smp_size)

train.t <- df.t[test_ind, ]
test.t <- df.t[-test_ind, ]
```

Random Forest
```{r, echo=FALSE}}
library(randomForest)
library(caret)

```
Fit the Random Forest Model
```{r, echo=FALSE}}
# Fit the model on the training set
set.seed(123)
model <- train(
  rating ~., data = train, method = "rf",
  trControl = trainControl("cv", number = 10),
  importance = TRUE
  )
# Best tuning parameter
model$bestTune

```

```{r, echo=FALSE}}
# Final model
model$finalModel
```
Find number of trees that produce lowest test MSE
```{r, echo=FALSE}}
#plot of the test MSE based on the number of trees used
plot(model$finalModel)
which.min(model$finalModel$mse)
#find RMSE of best model
sqrt(model$finalModel$mse[which.min(model$finalModel$mse)])


#From the output we can see that the model that produced the lowest test mean squared error (MSE) used 105 trees
```

Make predictions on the test data
```{r, echo=FALSE}}
predicted.classes <- model %>% predict(test)
head(predicted.classes)


pred_randomForest <- predict(model, test)
head(pred_randomForest)

```
Variable importance
```{r, echo=FALSE}}
importance(model)
```

```{r, echo=FALSE}}
varImp(model)
# Plot MeanDecreaseAccuracy
varImpPlot(model$finalModel, type = 1)
# Plot MeanDecreaseGini
varImpPlot(model$finalModel, type = 2)
```
Tune
This function produces the following plot, which displays the number of predictors used at each split when building the trees on the x-axis and the out-of-bag estimated error on the y-axis:
```{r, echo=FALSE}}

model_tuned <- tuneRF(
               x=train[,-1], #define predictor variables
               y=train$rating, #define response variable
               ntreeTry=500,
               mtryStart=4, 
               stepFactor=1.5,
               improve=0.01,
               trace=FALSE #don't show real-time progress
               )
```

Use the Final Model to Make Predictions

```{r, echo=FALSE}}
pred_s_train <- predict(model, newdata= train)
pred_s_test <- predict(model, newdata= test)
```
predicted values versus observed values in both training dataset and test dataset.
```{r, echo=FALSE}}
### 
plot(pred_s_train, train$rating)
abline(coef=c(0,1), col="red")
plot(pred_s_test, test$rating)
abline(coef=c(0,1), col="blue")

cor.test(pred_s_train, train$rating, method="pearson", conf.level=0.95)
cor.test(pred_s_test, test$rating, method="pearson", conf.level=0.95)
```

Hyperparameters
Note that, the random forest algorithm has a set of hyperparameters that should be tuned using cross-validation to avoid overfitting.
```{r, echo=FALSE}}

models <- list()
for (nodesize in c(1, 2, 4, 8)) {
    set.seed(123)
    model <- train(
      rating ~., data = train, method="rf", 
      trControl = trainControl(method="cv", number=10),
      nodesize = nodesize
      )
    model.name <- toString(nodesize)
    models[[model.name]] <- model
}
# Compare results
resamples(models) %>% summary()

```

#############
Computing KNN classifier
#
Fit the model on the training set
```{r, echo=FALSE}}
#trControl, to set up 10-fold cross validation
#preProcess, to normalize the data
#tuneLength, to specify the number of possible k values to evaluate
# Fit the model on the training set
set.seed(123)
model.knn <- train(
  rating ~., data = train, method = "knn",
  trControl = trainControl("cv", number = 10),
  preProcess = c("center","scale"),
  tuneLength = 10
  )
# Plot model accuracy vs different values of k
plot(model.knn)
```

Print the best tuning parameter k that
maximizes model accuracy
```{r, echo=FALSE}}
model.knn$bestTune
```

Make predictions on the test data
```{r, echo=FALSE}}
predicted.classes <- model.knn %>% predict(test)
head(predicted.classes) #knn#

```

Use the Final Model to Make Predictions

```{r, echo=FALSE}}
pred_knn_train <- predict(model.knn, newdata= train)
pred_knn_test <- predict(model.knn, newdata= test)
head(pred_knn_test)#knn

# results knn and rf - We see that our RMSE for our test is not very close to the RMSE we obtained on our model
caret::RMSE(pred_knn_train, train$rating)
caret::RMSE(pred_knn_train, test$rating)

#caret::RMSE(pred_s_train, train$rating)
caret::RMSE(pred_s_train, test$rating)


```
predicted values versus observed values in both training dataset and test dataset.
```{r, echo=FALSE}}
### 
plot(pred_knn_train, train$rating)
abline(coef=c(0,1), col="red")
plot(pred_knn_test, test$rating)
abline(coef=c(0,1), col="blue")

cor.test(pred_knn_train, train$rating, method="pearson", conf.level=0.95)
cor.test(pred_knn_test, test$rating, method="pearson", conf.level=0.95)

mean((pred_knn_train-train$rating)**2)
mean((pred_knn_test-test$rating)**2)
```


\textcolor{color}{For KNN Model , the correlation between predicted and observed rating is 0.3722099 in training dataset and 0.2718838  in test dataset, the 95\% CI of both don’t include 0. The ASE is 0.99 in training dataset and 1.03 in the test dataset. There is no overfitting  There is no overfitting.}

